---
title: "Diabetes Dashboard"
author: "Jiayi Zhang"
date: "2025-04-25"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(flexdashboard)
library(tidyverse)
library(plotly)
library(DT)
library(readr)  
library(rmarkdown)
options(scipen = 999)
```

Overview
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------
### Diabetes Prevalence by Age Group

```{r widget1, echo=FALSE, warning=FALSE, message=FALSE, results="asis"}
# read
file_path <- "C:/Users/grezh/Desktop/DATA555/archive/diabetes_012_health_indicators_BRFSS2015.csv"
df <- read_csv(file_path, show_col_types = FALSE)

# preprocessing
df_selected <- df %>%
  select(Diabetes_012, Age, BMI, PhysActivity, Smoker, HighBP, HighChol, Sex)

df_selected$Diabetes_012 <- factor(df_selected$Diabetes_012,
                                   levels = c(0, 1, 2),
                                   labels = c("No Diabetes", "Prediabetes", "Diabetes"))

df_selected$Age <- factor(df_selected$Age,
                          levels = 1:13,
                          labels = c("18-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54",
                                     "55-59", "60-64", "65-69", "70-74", "75-79", "80+"))

df_selected <- df_selected %>% filter(BMI <= 80)

# Prepare data
df_plot <- df_selected %>%
  group_by(Age, Diabetes_012) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(Age) %>%
  mutate(percentage = count / sum(count) * 100, total_n = sum(count))

# Plot
p1 <- ggplot(df_plot, aes(x = Age, y = percentage, fill = Diabetes_012,
                          text = paste0("Age: ", Age, 
                                        "<br>Diabetes: ", Diabetes_012,
                                        "<br>Percentage: ", round(percentage, 1), "%",
                                        "<br>Total N: ", total_n))) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("#1b9e77", "#ff7f00", "#7570b3")) +
  labs(title = "Diabetes Prevalence by Age Group",
       x = "Age Group", y = "Percentage (%)", fill = "Diabetes Status") +
  theme_minimal()

plotly1 <- ggplotly(p1, tooltip = "text")
plotly1
```

### ðŸ”¹ Key Takeaways

- Diabetes prevalence increases significantly with age.
- Individuals aged 60 years and older have the highest diabetes rates.
- Prediabetes remains relatively low across all age groups, which may indicate underdiagnosis or underreporting.

Column {data-width=350}
-----------------------------------------------------------------------

### BMI Distribution by Diabetes Status

```{r widget2, echo=FALSE, warning=FALSE, message=FALSE, results="asis"}
# box plot
p2 <- ggplot(df_selected, aes(x = Diabetes_012, y = BMI, fill = Diabetes_012)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA) +
  stat_summary(fun = mean, geom = "point", shape = 4, size = 3, color = "black") +  # æ·»åŠ å‡å€¼ (X)
  stat_summary(fun = median, geom = "point", shape = 3, size = 3, color = "red") +   # æ·»åŠ ä¸­ä½æ•° (+)
  labs(title = "BMI Distribution by Diabetes Status",
       x = "Diabetes Status", y = "BMI (kg/mÂ²)", fill = "Diabetes Status") +
  scale_fill_manual(values = c("#1b9e77", "#ff7f00", "#7570b3")) +
  theme_minimal()

plotly2 <- ggplotly(p2, tooltip = "y")
plotly2
```

### ðŸ”¹ Key Takeaways

- Individuals with diabetes tend to have higher median BMI than those without diabetes.
- Higher BMI values are strongly associated with an increased risk of diabetes.
- Outliers with extreme BMI (>80) were removed to enhance visualization clarity.

### Interactive Data Table

```{r widget3, echo=FALSE, warning=FALSE, message=FALSE, results="asis"}
datatable(df_selected,
          options = list(pageLength = 10, autoWidth = TRUE, searchHighlight = TRUE),
          filter = "top", rownames = FALSE)
```

### ðŸ”¹ Key Takeaways

- The table allows users to search, filter, and explore diabetes health indicators interactively.
- Filtering for higher BMI values (e.g., BMI > 30) can reveal potential patterns in diabetes prevalence.

Dataset
=======================================================================

### Dataset Description

- **Source:** [Diabetes Health Indicators Dataset on Kaggle](https://www.kaggle.com/datasets/alexteboul/diabetes-health-indicators-dataset)
- **Sample Size:** 253,680 records
- **Data Collection Method:** Survey responses collected by the Behavioral Risk Factor Surveillance System (BRFSS) in 2015.
- **Study Population:** Adults aged 18 years and older living in the United States.
- **Time Period:** Data collected during the 2015 calendar year.

About
=======================================================================

### Project Importance

This dashboard provides interactive insights into key risk factors for diabetes.  
It helps public health practitioners design targeted interventions to reduce diabetes prevalence.

---

### Source Code Repository

You can view the full source code for this dashboard on GitHub:  
[GitHub Repository Link](https://github.com/Jiayi-618/diabetes_dashboard)

